### 2025-08-21 base-rates.qmd
20250821 - Bayesian updating of fantasy points

@@

Using [Bayes' theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
The calculation of posttest (posterior) probability is described in @sec-posttestProbability.
A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
The updated (posttest) probability is where the connecting line crosses the third, right line.

>>>>

Bayesian updating can also be applied to continuous outcomes like fantasy points.
For an example of applying Bayesian updating to fantasy points, see @Braun2012 (archived at <https://web.archive.org/web/20161028142225/http://www.bayesff.com/bayesian101/>).
Applying Bayes' theorem to continuous outcomes, the posterior distribution is equal to the prior distribution times the likelihood (data).
For instance, we can start with our prior belief (distribution) for a player's performance based on, for example, average draft position.
Then, we observe the Week 1 performance.
For instance, if Tom Brady scores 35 points in Week 1, our likelihood represents the likelihood that Tom Brady would score 35 points (the data).
Using Bayesian updating, we can then calculate a posterior distribution that represents our new best prediction moving forward for how many points Tom Brady will score in Week 2.
We then observe Week 2, generate a new likelihood and posterior distribution, and use that as our new prior distribution for Week 3, and so on.
To keep things simple, we use a binary outcome below to demonstrate Bayesian updating.

@@

Using [Bayes' theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
The calculation of posttest (posterior) probability is described in @sec-posttestProbability.
In its calculation, the probability nomogram automatically converts the pretest probability (i.e., base rate) to prior (pretest) odds and the posterior (posttest) odds to posterior probability, so you do not have to.
A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
The updated (posttest) probability is where the connecting line crosses the third, right line.
